---
title: "Development Code"
output: pdf_document
---

This file was used to create two data frames, 'by_match' and 'by_country', that were then primarily used in the data_analysis.Rmd file.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(StatsBombR)
library(ggplot2)
library(grid)
library('writexl')
library('readxl')
library('xlsx')
```

# Reading data in
```{r}
Comp <- FreeCompetitions()
Matches <- FreeMatches(Comp)
Matches = Matches %>% filter(competition.competition_name=="UEFA Euro")
data360 <- StatsBombFree360Events(MatchesDF = Matches, Parallel = T)

events <- StatsBombFreeEvents(MatchesDF = Matches, Parallel = T)
events <- allclean(events)
events <- get.opposingteam(events)

data360 = data360 %>% rename(id = event_uuid)

events = events %>% left_join(data360, by = c("id" = "id"))

events = events %>% rename(match_id = match_id.x) %>% select(-match_id.y)

```

# Making by team data  
Creating function for each team
```{r}
new_row_country <- function(team_name){
  
  duels = events %>%
  filter(team.name == toString(team_name), type.id == 4, duel.type.id != 10, freeze_frame != 'NULL') %>%
  select(id, team.name, OpposingTeam, player.name, minute, second, location.x, location.y, freeze_frame, duel.outcome.id)


num_dist = 5
team_close_vect <- c()
team_far_vect <- c()

opp_close_vect <- c()
opp_far_vect <- c()

team_dist_vect <- c()
opp_dist_vect <- c()

diff_close_vect <- c()
diff_far_vect <- c()

for (n in 1:dim(duels)[1]) {
  
  duelone = duels[n,] %>% unnest(freeze_frame) %>%
mutate(ff_location.x = (map(location, 1)), ff_location.y = (map(location, 2))) %>%
select(-location) %>%
mutate(ff_location.x = as.numeric(ifelse(ff_location.x == 'NULL', NA, ff_location.x)), ff_location.y = as.numeric(ifelse(ff_location.y == 'NULL', NA, ff_location.y))) 
  
  
  # get one duel' actor and everyoneelse frames
  actor = duelone %>%
    filter(actor == TRUE)
  
  teammates = duelone %>%
    filter(actor == FALSE, teammate == TRUE, keeper == FALSE)
  
  opp = duelone %>%
    filter(actor == FALSE, teammate == FALSE, keeper == FALSE)

  # get # of players who are far or near in 
  team_far = 0
  team_close = 0
  for (i in 1:dim(teammates)[1]){
    distance = sqrt((as.numeric(actor[1, 13])-as.numeric(teammates[i,13]))^2 + (as.numeric(actor[1, 14])-as.numeric(teammates[i,14]))^2)
    
    
    if (is.na(distance)){distance <- 0}
    team_dist_vect = append(team_dist_vect, distance)
    
    if (distance > num_dist){team_far = team_far + 1}
    else if (distance > 0 && distance < num_dist) {team_close = team_close + 1}
  }
  
  team_far_vect <- append(team_far_vect, team_far)
  team_close_vect <- append(team_close_vect, team_close)
  
  opp_far = 0
  opp_close = 0
  for (i in 1:dim(opp)[1]){
    distance = sqrt((as.numeric(actor[1, 13])-as.numeric(opp[i,13]))^2 + (as.numeric(actor[1, 14])-as.numeric(opp[i,14]))^2)
    
    
    if (is.na(distance)){distance <- 0}
    opp_dist_vect = append(opp_dist_vect, distance)
    
    if (distance > num_dist){opp_far = opp_far + 1}
    else if (distance > 0 && distance < num_dist) {opp_close = opp_close + 1}
  }
  
  opp_far_vect <- append(opp_far_vect, opp_far)
  opp_close_vect <- append(opp_close_vect, opp_close)
  
  diff_far = team_far - opp_far
  diff_close = team_close - opp_close
  
  diff_close_vect <- append(diff_close_vect, diff_close)
  diff_far_vect <- append(diff_far_vect, diff_far)
  
}


duels2 = duels 
duels2['team_close'] <- team_close_vect
duels2['team_far'] <- team_far_vect

duels2['opp_close'] <- opp_close_vect
duels2['opp_far'] <- opp_far_vect

duels2['diff_close'] <- diff_close_vect
duels2['diff_far'] <- diff_far_vect


duels3 = duels2 %>% 
  mutate(duel.outcome = ifelse(duel.outcome.id == 14 | duel.outcome.id == 13,"lost", "won"))

duels_0_total = duels3 %>%
  filter(diff_close == 0)
duels_0_won = duels3 %>%
  filter(diff_close == 0, duel.outcome == "won")

duels_1_total = duels3 %>%
  filter(diff_close == 1)
duels_1_won = duels3 %>%
  filter(diff_close == 1, duel.outcome == "won")

duels_2_total = duels3 %>%
  filter(diff_close == -1)
duels_2_won = duels3 %>%
  filter(diff_close == -1, duel.outcome == "won")

duels_3_total = duels3 %>%
  filter(diff_close == 2)
duels_3_won = duels3 %>%
  filter(diff_close == 2, duel.outcome == "won")

duels_4_total = duels3 %>%
  filter(diff_close == -2)
duels_4_won = duels3 %>%
  filter(diff_close == -2, duel.outcome == "won")



min2 <- c(nrow(duels_4_won)/nrow(duels_4_total)) # -2

plus2 <- c(nrow(duels_3_won)/nrow(duels_3_total)) # 2

zero <- c(nrow(duels_0_won)/nrow(duels_0_total)) # 0

plus1 <- c(nrow(duels_1_won)/nrow(duels_1_total)) # 1

min1 <- c(nrow(duels_2_won)/nrow(duels_2_total)) # -1

df <- data.frame(team_name, min2, min1, zero, plus1, plus2)

return(df)}
```

Looping over every team name
```{r}
teams1 = events %>%
  select(team.name)

teams = unique(teams1)

df_by_team <- data.frame(TeamName = character(), Down2 = double(), Down1 = double(), Equal = double(), Up1 = double(), Up2 = double(), stringsAsFactors=FALSE)


for (i in 1:dim(teams)[1]) {
  nameofteam = toString(teams[i,1]) 
  arow = new_row_country(nameofteam)
  df_by_team <- rbind(df_by_team, arow)}

df_by_team
```

Joining other spreadsheet with team stats
```{r}
stats_by_country <- read.csv("extra_data_bycountry.csv")

by_country = stats_by_country %>% right_join(df_by_team, by = c("team_name" = "team_name"))
```



# Making by match data
Creating function for each match
```{r}
new_row_match <- function(match, team_name){
  
  duels = events %>%
  filter(match_id == match, team.name == toString(team_name), type.id == 4, duel.type.id != 10, freeze_frame != 'NULL') %>%
  select(id, team.name, OpposingTeam, player.name, minute, second, location.x, location.y, freeze_frame, duel.outcome.id, match_id)


match = duels[1,11]

num_dist = 5
team_close_vect <- c()
team_far_vect <- c()

opp_close_vect <- c()
opp_far_vect <- c()

team_dist_vect <- c()
opp_dist_vect <- c()

diff_close_vect <- c()
diff_far_vect <- c()

for (n in 1:dim(duels)[1]) {
  
  duelone = duels[n,] %>% unnest(freeze_frame) %>%
mutate(ff_location.x = (map(location, 1)), ff_location.y = (map(location, 2))) %>%
select(-location) %>%
mutate(ff_location.x = as.numeric(ifelse(ff_location.x == 'NULL', NA, ff_location.x)), ff_location.y = as.numeric(ifelse(ff_location.y == 'NULL', NA, ff_location.y))) 
  
  
  # get one duel' actor and everyoneelse frames
  actor = duelone %>%
    filter(actor == TRUE)
  
  teammates = duelone %>%
    filter(actor == FALSE, teammate == TRUE, keeper == FALSE)
  
  opp = duelone %>%
    filter(actor == FALSE, teammate == FALSE, keeper == FALSE)

  # get # of players who are far or near in 
  team_far = 0
  team_close = 0
  for (i in 1:dim(teammates)[1]){
    distance = sqrt((as.numeric(actor[1, 13])-as.numeric(teammates[i,13]))^2 + (as.numeric(actor[1, 14])-as.numeric(teammates[i,14]))^2)
    
    
    if (is.na(distance)){distance <- 0}
    team_dist_vect = append(team_dist_vect, distance)
    
    if (distance > num_dist){team_far = team_far + 1}
    else if (distance > 0 && distance < num_dist) {team_close = team_close + 1}
  }
  
  team_far_vect <- append(team_far_vect, team_far)
  team_close_vect <- append(team_close_vect, team_close)
  
  opp_far = 0
  opp_close = 0
  for (i in 1:dim(opp)[1]){
    distance = sqrt((as.numeric(actor[1, 13])-as.numeric(opp[i,13]))^2 + (as.numeric(actor[1, 14])-as.numeric(opp[i,14]))^2)
    
    
    if (is.na(distance)){distance <- 0}
    opp_dist_vect = append(opp_dist_vect, distance)
    
    if (distance > num_dist){opp_far = opp_far + 1}
    else if (distance > 0 && distance < num_dist) {opp_close = opp_close + 1}
  }
  
  opp_far_vect <- append(opp_far_vect, opp_far)
  opp_close_vect <- append(opp_close_vect, opp_close)
  
  diff_far = team_far - opp_far
  diff_close = team_close - opp_close
  
  diff_close_vect <- append(diff_close_vect, diff_close)
  diff_far_vect <- append(diff_far_vect, diff_far)
  
}


duels2 = duels 
duels2['team_close'] <- team_close_vect
duels2['team_far'] <- team_far_vect # num of players > 10 of actor
# dist_vect

duels2['opp_close'] <- opp_close_vect
duels2['opp_far'] <- opp_far_vect

duels2['diff_close'] <- diff_close_vect
duels2['diff_far'] <- diff_far_vect


duels3 = duels2 %>% 
  mutate(duel.outcome = ifelse(duel.outcome.id == 14 | duel.outcome.id == 13,"lost", "won"))

duels_0_total = duels3 %>%
  filter(diff_close == 0)
duels_0_won = duels3 %>%
  filter(diff_close == 0, duel.outcome == "won")

duels_1_total = duels3 %>%
  filter(diff_close == 1)
duels_1_won = duels3 %>%
  filter(diff_close == 1, duel.outcome == "won")

duels_2_total = duels3 %>%
  filter(diff_close == -1)
duels_2_won = duels3 %>%
  filter(diff_close == -1, duel.outcome == "won")

duels_3_total = duels3 %>%
  filter(diff_close == 2)
duels_3_won = duels3 %>%
  filter(diff_close == 2, duel.outcome == "won")

duels_4_total = duels3 %>%
  filter(diff_close == -2)
duels_4_won = duels3 %>%
  filter(diff_close == -2, duel.outcome == "won")



min2 <- c(nrow(duels_4_won)/nrow(duels_4_total)) # -2

plus2 <- c(nrow(duels_3_won)/nrow(duels_3_total)) # 2

zero <- c(nrow(duels_0_won)/nrow(duels_0_total)) # 0

plus1 <- c(nrow(duels_1_won)/nrow(duels_1_total)) # 1

min1 <- c(nrow(duels_2_won)/nrow(duels_2_total)) # -1

df <- data.frame(match, team_name, min2, min1, zero, plus1, plus2)

return(df)}
```

Looping over every match and team name combination
```{r}
all_matches1 = events %>%
  select(match_id, team.name, OpposingTeam)

all_matches = unique(all_matches1)

df_by_match <- data.frame(MatchID = integer(), TeamName = character(), Down2 = double(), Down1 = double(), Equal = double(), Up1 = double(), Up2 = double(), stringsAsFactors=FALSE)

x = all_matches

for (i in 1:dim(x)[1]) {
team = x[i, 'team.name']
match = x[i, 'match_id']
arow = new_row_match(as.integer(match), toString(team))
df_by_match <- rbind(df_by_match, arow)
}

df_by_match
```

Joining other spreadsheet with match stats (created from Python file)
```{r}
match_stats = read.csv("bymatch_stats.csv")

by_match <- merge(df_by_match, match_stats,by=c('match_id',"team_name"))
```
